---
output: html_document
---
# Reproducible Research: Assignment 1
Adam Suhajda, 20160902
========================================================

## Introducing task and getting data

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
Create a new dataset that is equal to the original dataset but with the missing data filled in.
Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
library(lubridate)
setwd("C:/Users/suhaada/Desktop/R/Coursera/Reproducible res")
dataset <- read.csv("./activity.csv")
dataset$data <- as.Date(dataset$date, format="%Y-%m-%d")
dataset$weekday <- wday(dataset$date, label=T,abbr=T)
dataset$weekday <- as.factor(dataset$weekday)
dataset$interval <- as.factor(dataset$interval)
days <- as.factor(dataset$date)
```

## Question 1:total number of steps taken per day

```{r}
total <- aggregate(steps ~ date, dataset, sum)
hist(total$steps, xlab= "Total steps per day", col="red", breaks = 10)
```

mean of steps taken per day

```{r}
mean(total$steps)
```

median of steps taken per day

```{r}
median(total$steps)
```

## Question 2:  average of daily activity pattern
```{r}
listaIntervalli <- split(x=dataset,f=dataset$date)
listaIntervalli <- listaIntervalli[[1]]['interval']



library(lattice)
mediaStep <- tapply(dataset$steps, dataset$interval, mean, na.rm=T )


xyplot(mediaStep ~ listaIntervalli,
       type="l",
       ylab="mean steps",
       xlab="intervals",
       las=2,
       par.settings=list(layout.heights=list(top.padding=3, bottom.padding=5)))
```

Which 5-minute interval contains the maximum number of steps?

```{r}
names(which.max(x=mediaStep))
```

## Question 3: Imputing missing values

 number of missing values
```{r}
average <- tapply(dataset$steps, dataset$weekday, mean, na.rm=T )
naElem <- dataset[is.na(dataset$steps),]
length(naElem$steps)   
```



Exchange null values to mean values
```{r}

days <- as.factor(dataset$date)
X <- split(dataset, days)

for(i in 1:length(X))
{
  for(j in 1:length(X[[i]][,'steps']))
  {
    
    if(is.na(X[[i]][j,'steps']))
    {
      
      giorno <- X[[i]][j,'date']
      
      X[[i]][j,'steps'] <- average[wday(giorno)]
    }  
  }
  
}

# refreshing dataset environmental variable
newdataset <- data.frame(steps={},
                         date={},
                         interval={})
for(i in 1:length(X))
{
  newdataset <- rbind(newdataset, data.frame(steps=X[[i]]['steps'],
                                             date=X[[i]]['date'],
                                             interval=X[[i]]['interval']))
  
}

```

mean value rerun
```{r}
tapply(newdataset$steps, dataset$weekday, mean, na.rm=T )

```

median value rerun
```{r}
tapply(newdataset$steps, dataset$weekday, median, na.rm=T )
```

histogram 
```{r}
total_new <- aggregate(steps ~ date, newdataset, sum)
hist(total_new$steps, xlab= "Total steps per day", main="", col="blue", breaks = 10)
```


## Question 4:  differences in activity patterns on weekdays and weekends

```{r}
# refreshing dataset environmental variable
newdataset['dayTypeInWeek'] = factor(sapply(newdataset$date, function(x){ if (wday(x) == 6 | wday(x) == 7) { "weekend" } else { "weekday"} }))


avgStepdayTypeInWeek = aggregate(steps~interval + dayTypeInWeek, mean, data=newdataset)

library(lattice)
xyplot( steps ~ interval | dayTypeInWeek, data = avgStepdayTypeInWeek, type="l", layout=c(1,2), xlab="Interval", ylab="Number of steps")
```
